{\includegraphics[width=3cm, height=2.5cm]{Cran.jpg}} \\ \includegraphics[width=7cm, height=1cm]{TB.jpg} \\ {Technical Documentation for TersectBrowser+}

### 2025-05-05 TechDoc/TechDoc.tex
Adjust architecture section

The frontend is a browser-based interface developed in Angular and React, designed for interactive data visualization, including distance plots, accession selectors, and genome browser integration. The backend, built in Node.js and Express, provides API endpoints, data models, and utility scripts to serve genomic data and computation results from a MongoDB database. A set of common modules supports both frontend and backend logic with shared resources. Additionally, the architecture supports pluggable extensions, such as the genome-browser, which provide domain-specific capabilities like rendering gene models or enabling barcode-based comparisons. Supporting data files, environment configurations, and Python scripting for preprocessing or bioinformatics workflows are organized in separate folders for data loading and virtual environments, enabling efficient pipeline integration and deployment.

@@

\textit{Tersect Browser+} is primarily built using Angular 8, preserving the architecture of the original application. To enable modern feature development without disrupting the existing codebase, a dedicated extension layer has been introduced using Angular 15 and Node.js 22. This modular, "bolt-on" extension framework allows for the seamless integration of new capabilities while taking advantage of the latest frontend and backend technologies. Extensions are developed independently and published to NPM, making them easily linkable through project dependencies. This hybrid approach ensures forward compatibility and scalability, while maintaining stability in the core application. Future versions of \textit{Tersect Browser+} may explore fully migrating the main application to a more unified and modern framework.

@@

\subsubsection{Test Results}
This extension has been tested with the additional Soybean dataset, with correct functionality. The relevance of the gene models in the GFF file depend on what the user provides in the dataset, and how the GFF file was created.

\subsection{Feature Search Extension}
\subsubsection{Rationale}
The aim of this extension is to allow the user to search selected intervals and identify which accessions contain \hl{high/medium/low} impact variants within that region. The bins containing high impact variants for that gene will be highlighted at the accession level, allowing high resolution investigation of variants. The highlighted bins may then be selected and opened in the popup \nameref{sec:Browser} window for further investigation.

@@

\subsection*{4.3.3 Feature Search Extension – Technical Description}

\subsubsection*{4.3.3.1 Creating Search Indexes Per Chromosome}

@@

The final VCF was then compressed using \texttt{bgzip} and indexed with \texttt{tabix}. Additionally, Trix index files \citep{ucsc_trix} were generated for high impact variants on each chromosome to support rapid gene name search suggestions.

\subsubsection*{4.3.3.2 Handling Search Request}

@@

\subsubsection*{4.3.3.3 Handling Searching by Interval}

@@

\subsubsection*{4.3.3.4 Handling Searching by Gene Name}

@@

\subsubsection*{4.3.3.5 Linking with the Frontend}

@@

\subsubsection*{4.3.3.6 Other Ways to Trigger Variant Search}

@@

\subsubsection*{4.3.3.7 Highlighting Bins}

@@

\subsubsection*{VCF Mechanism Summary}

>>>>

The frontend is a browser-based interface developed in Angular JS designed for interactive data visualization, including distance plots, accession selectors.

 The backend, built in Node.js and Express, provides API endpoints, data models, and utility scripts to serve genomic data and computation results from a MongoDB database.

 A set of common modules supports both frontend and backend logic with shared resources.  Additionally, the architecture supports pluggable extensions, such as the genome-browser which provide domain-specific capabilities like rendering gene models or enabling barcode-based comparisons.

 Supporting data files, environment configurations, and Python scripting for preprocessing or bioinformatics workflows are organized in separate folders for data loading and virtual environments, enabling efficient pipeline integration and deployment.

\subsection{Embedding The Extensions }

@@

\textit{Tersect Browser+} is originally built using Angular 8, to enable modern feature development without disrupting the existing codebase, a dedicated extension layer has been introduced using Angular 15 and Node.js 22.

New extensions are converted to web components and bundled with rollup [TODO: Reference] to a single min.js bundle, the target bundle is then transpiled to a version compatible with older browsers with babel [TODO: Reference]. The final output is deployed on NPM.

The genomebrowser relies on \texttt{Jbrowse} [TODO: Reference] extensively. Since Jbrowse is written in \texttt{React} [TODO: Reference], it is embedded in the new angular app via the \texttt{ReactComponentDirective} provided by the \texttt{@ngeenx/ngx-react} package on NPM.

This modular,” bolt-on” extension framework allows for the seamless integration of new capabilities while taking advantage of the latest frontend technologies.

Extensions are developed independently and published to NPM, making them easily linkable through project dependencies. This hybrid approach ensures forward compatibility and scalability, while maintaining stability in the core application. Future versions of Tersect Browser+ may explore fully migrating the main application to a more unified and modern framework.

@@

\subsubsection{Test Results}
This extension has been tested with the additional Soybean dataset, with correct functionality. The relevance of the gene models in the GFF file depend on what the user provides in the dataset, and how the GFF file was created. 
\clearpage
\subsection{Feature Search Extension}
\subsubsection{Rationale}
The aim of this extension is to allow the user to search selected intervals and identify which accessions contain \hl{high/medium/low} impact variants within that region. The bins containing high impact variants for that gene will be highlighted at the accession level, allowing high resolution investigation of variants. The highlighted bins may then be selected and opened in the popup \nameref{sec:Browser} window for further investigation.

@@

\subsection{Feature Search Extension - Technical Description}

\subsubsection{Creating Search Indexes Per Chromosome}

@@

The final VCF was then compressed using \texttt{bgzip} and indexed with \texttt{tabix}. Additionally, Trix index files [TODO:reference] were generated for high impact variants on each chromosome to support rapid gene name search suggestions.

\subsubsection{Handling Search Request}

@@

\subsubsection{Handling Searching by Interval}

@@

\subsubsection{Handling Searching by Gene Name}

@@

\subsubsection{Linking with the Frontend}

@@

\subsubsection{Other Ways to Trigger Variant Search}

@@

\subsubsection{Highlighting Bins}

@@

\subsubsection{VCF Mechanism Summary}
